---
title: "week 1"
output: html_notebook
---

Newtown-Raphton algorithm to find an estimate of the Weibull scale parameter (theta)

```{r}
pressure <- c(1051, 1337, 1389, 1921, 1942, 2322, 3629, 4006, 4012,
              4063, 4921, 5445, 5620, 5817, 5905, 5956, 6068, 6121,
              6473, 7501, 7886, 8108, 8546, 8666, 8831, 9106, 9711,
              9806, 10205, 10396, 10861, 11026, 11214, 11362, 11604,
              11608, 11745, 11762, 11895, 12044, 13520, 13670, 14110,
              14496, 15395, 16179, 17092, 17568, 17568)

hist(pressure, nclass=17, xlim=c(0,20000), xlab="Time to failure (hours)")

## Maximum likelihood estimator of the Weibull scale parameter (theta).
# The shape parameter (lambda) is fixed.

theta.hat <- function(y, lambda){
  ( sum(y^lambda) / length(y) )^(1/lambda)
}

theta.hat(y=pressure, lambda=2)
```

Write some RR code to reproduce this plot:

insert picture

```{r}
pressure <- c(1051, 1337, 1389, 1921, 1942, 2322, 3629, 4006, 4012,
              4063, 4921, 5445, 5620, 5817, 5905, 5956, 6068, 6121,
              6473, 7501, 7886, 8108, 8546, 8666, 8831, 9106, 9711,
              9806, 10205, 10396, 10861, 11026, 11214, 11362, 11604,
              11608, 11745, 11762, 11895, 12044, 13520, 13670, 14110,
              14496, 15395, 16179, 17092, 17568, 17568)

hist(pressure, nclass=17, xlim=c(0,20000), xlab="Time to failure (hours)")
dev.copy2pdf(file="lect2_pressures.pdf", height=4, width=8)

## Maximum likelihood estimator of the Weibull scale parameter (theta)
## The shape parameter (lambda) is fixed.

theta.hat <- function(y, lambda){
  ( sum(y^lambda) / length(y) )^(1/lambda)
}
```

# 1-5
Draw the probability mass function of a Binomial random variable with $n=40$ and $p=0.5$ in R.

my attempt
```{r}
n <- 0:40
plot(n, dbinom(n, size=40, prob=.5))
```

solution
```{r}
# Binomial distribution
# Probability mass function of a Binomial random variable with n=40 and p=0.5
xseq <- 0:40
plot(xseq, dbinom(xseq, size=40, prob=0.5), ylab=expression(f(x)), pch=16)
```

Draw the probability density function of a Normal random variable with μ=4 and σ=1.5 in R.

```{r}

```

Draw the probability density function of a Poisson random variable with $\lambda = 1$ in R.

my attempt
```{r}

```

solution
```{r}
## Normal distribution
# Probability density function of a Normal random variable with mu=4 and sigma=1.5
xseq <- seq(from=-10, to=10, length=100)
plot(xseq, dnorm(xseq, mean=4, sd=1.5), type="l", ylab=expression(f(x)))
```


### 1-5-3
Draw the probability density function of a Poisson random variable with λ=1 in R.
```{r}
# Probability density function of a Poisson random variable with mu=4 and sigma=1.5
xseq <- 0:10
plot(xseq, dpois(xseq, lambda=1), ylab=expression(f(x)), pch=16)
```


### 1-5 question 1
Use RR to complete the following tasks:

Draw the probability mass function of a Binomial random variable with n=40 and p=0.5

```{r}
# Probaility mass function of a Binomial random variable with n=40 and p=0.5
xseq <- 0:40
plot(xseq, dbinom(xseq, size=40, prob=0.5), ylab=expression(f(x)), pch=16)
```

Draw the probability density function of a Normal random variable with μ=4 and σ=1.5

```{r}
# Probaility density function of a Normal random variable with mu=4 and sigma=1.5
xseq <- seq(from=-10, to=10, length=100)
plot(xseq, dnorm(xseq, mean=4, sd=1.5), type="l", ylab=expression(f(x)))
```

Draw the probability density function of a Normal random variable with μ=4μ=4 and σ=1.5σ=1.5.

```{r}
# Probaility density function of a Normal random variable with mu=4 and sigma=1.5
xseq <- seq(from=-10, to=10, length=100)
plot(xseq, dnorm(xseq, mean=4, sd=1.5), type="l", ylab=expression(f(x)))
```

Draw the probability density function of a Poisson random variable with λ=1.

```{r}
# Probaility density function of a Poisson random variable with mu=4 and sigma=1.5
xseq <- 0:10
plot(xseq, dpois(xseq, lambda=1), ylab=expression(f(x)), pch=16)
```



### 1-5 question 2

```{r}
```



