---
title: "model selection swiss dataset"
output: html_notebook
---

# Activity in R: Model Selection
Consider the swiss dataset in R including standardised fertility measure and
socio-economic indicators for each of 47 French-speaking provinces of
Switzerland at about 1888 head(swiss)
             Fertility  Agriculture  Examination  Education  Catholic
Courtelary        80.2         17.0           15         12      9.96
Delemont          83.1         45.1            6          9     84.84
Franches-Mnt      92.5         39.7            5          5     93.40
Moutier           85.8         36.5           12          7     33.77
Neuveville        76.9         43.5           17         15      5.16
Porrentruy        76.1         35.3            9          7     90.57
             Infant.Mortality
Courtelary               22.2
Delemont                 22.2
Franches-Mnt             20.2
Moutier                  20.3
Neuveville               20.6
Porrentruy               26.6

Calculate AIC and BIC for the multiple regression model with Fertility as a
response and all other variables as explanatory variables.

Additionally, consider a model with ExaminationExamination removed from the
set of the explanatory variables. Compare the AIC and BIC for the simplified
model and the full model.

Hint: Use functions `AIC()` and `BIC()` for this exercise.

# solution

```{r}
library(datasets) # Load the datasets package, if not already loaded
data(swiss)  # Access the Swiss dataset
attach(swiss)  # Attach the dataset to the search path
head(swiss)  # View the first few rows of the dataset
```

```{r}
lm1 <- lm(Fertility ~ . , data = swiss)
lm2 <- update(lm1, . ~ . -Examination)

AIC(lm1, lm2)
##     df      AIC
## lm1  7 326.0716
## lm2  6 325.2408
```



```{r}
BIC(lm1, lm2)
##     df      BIC
## lm1  7 339.0226
## lm2  6 336.3417
```



```{r}
head(swiss)

lm1 <- lm(Fertility ~ . , data = swiss)
lm2 <- update(lm1, . ~ . -Examination)

AIC(lm1, lm2)

```



```{r}
BIC(lm1, lm2)
```