---
title: "Loess Smoother"
output: html_notebook
---

# Activity in R: Loess Smoother
Download the data set ethanol. The ethanol data frame contains 8 measurements generated in an experiment in which ethanol was burned in a single cylinder automobile test engine. The variables are `NOx` (a measure of the nitrogen oxides in the exhaust), `C` (compression ratio of an engine) and `E` (a measure of the richness of the air/ethanol mix used).

```{r}
ethanol<-read.table("data/ethanol.txt",header=T)
attach(ethanol)
```

```{r}
print(dim(ethanol))
```

```{r}
head(ethanol)
```

(a) Do a scatterplot of `NOx` against `E`.
After looking at the scatterplot you might suggest a polynomial regression model to capture the relationship between the response NOxNOx and the predictor `E`. Fit a quadratic regression model and superimpose the fitted curve on the scatterplot.

```{r}
plot(E, NOx)
Esq<-E*E
ethanol.lm<-lm(NOx~E+Esq)
Egrid<-seq(from=min(E),to=max(E),length=50)

new.data<-data.frame(E=Egrid,Esq=Egrid*Egrid)
ethanol.pred<-predict(ethanol.lm,newdata=new.data)
lines(Egrid,ethanol.pred,col=2)
lines(loess.smooth(E,NOx,span=0.2),lty=2,col=3)
```

(b) Superimpose on the scatterplot in a) a suitable loess smooth. Do you think the smooth captures structure in the data that is missed by the quadratic regression model?

```{r}

```
